### 2014-01-27 brains-protocol.tex
Code style
3.3 is reread and good.

In order to start bridging this gap, we set out to \emph{empirically} study public representation transformations at the microscopic level, aiming to stay compatible with macroscopic-level studies of these public representations.
Quotations appeared to be a perfect candidate as public representations.
First, they are usually cleanly delimited by quotation marks (and often with HTML markup in web pages), which greatly facilitates their detection in text corpora.
Second, they stem from a unique ``original'' version, and could ideally be traceable back to that version.
Third, and most importantly, their duplication should \emph{a priori} be highly faithful, apart from cases of cropping: not only should transformations be of moderate magnitude, but when specific words are not perfectly duplicated, it is safe to assume that the variation is due to involuntary cognitive bias --- as writers may expect any casual reader to easily verify, and thus criticize, the fidelity to the original quote.
Quotation evolution is therefore a perfect environment to measure cognition-induced transformations and relate those findings to macroscopic social dynamics.

@@

Nonetheless, these sets bear no explicit mention of which is the authoritative, original quotation.  Defining substitutions from the dataset therefore requires additional assumptions.  
%\tb{Il manque en effet certaines informations dans le corpus pour pouvoir lire ces substitutions de façon immédiate dans les données : les métadonnées pour chaque citation ne concernant que les URLs, les dates d'apparitions et les nombres d'occurrences à chaque URL, on n'a aucune information sur les liens entre les citations. Or, on a besoin de savoir sur quelles sources un auteur s'est appuyé pour reproduire une citation, c'est-à-dire qu'on a besoin de savoir quelle citation a été générée à partir de quelle autre. Cette information n'est pas présente dans les données et on va donc devoir l'inférer, à l'aide de plusieurs hypothèses.}

Indeed, an author writing a new post and reproducing (possibly altering) a quotation bases himself on original quotations we have no access to. We face a reverse engineering problem where, given all quotations and their occurrence timestamps, we must estimate which was the originating quote for each instance of each quote. We therefore model the underlying quote selection process by making a few additional assumptions which let us define substitutions from the available data.

The main question in this problem concerns the way we consider later occurrences of a quote which, when it first appeared, was identified to be an alteration from an original quote. Let us give an example: say the quotation ``These accusations are false and absurd'' ($q_a$) appears in a blog on Jan 18th, and the slightly different quotation ``These accusations are false and incoherent'' ($q_b$) appears in another blog on the 19th, the 20th, and the 21st of January. If $q_a$ was sufficiently prominent when $q_b$ first appeared, we can safely assume that the author of $q_b$ on the 19th based himself on $q_a$. But what about the following occurrences of $q_b$? Should we consider them to be substitutions based on $q_a$ (i.e. re-creations of $q_b$ by a new instance of the substitution process that brought from $q_a$ to $q_b$ in the first place) or reproductions of the first occurrence of $q_b$?

To settle this question we model the process as follows: we assume that when a quote $q$ appears at time $t$, if not original (i.e. if not stemming from a source external to the dataset, e.g. initiating a new set of quotations), it is based solely on the most frequent quote $q_{max}$ in the preceding period of time $[t - \Delta t ; t]$. The length $\Delta t$ of that period of time is fixed as a fraction of the total duration of the considered set of quotations; we took one fifth in the implementation (i.e. one month for a set of quotations spanning five months, or one day for a set spanning only five days; this takes the dynamism of the set of quotes into account). If $q$ differs from $q_{max}$ by only a word, it is counted as a substitution from $q_{max}$ to $q$. In any other case, i.e. if $q$ and $q_{max}$ are the same or if $q$ and $q_{max}$ are different in other ways, the occurrence of $q$ is not considered to be an instance of substitution and is discarded.

@@

The assumptions embedded in this model are only a subset of a wider set of possibilities, all leading to alternative models. We identified and implemented five other such possible models, which all yielded essentially the same results. These models differ in the definition of the source quote from which new occurrences stem, essentially modifying the balance between reproduction of previous occurrences of a quote and re-creation of itself by a new substitution instance. For example, the time-windows considered can have different lengths, can include all occurrences from the beginning of the set of quotations, or can have fixed positions. The source quote of a potential substitution can be chosen among those time-windows, or otherwise (e.g. among \emph{all} quotes having appeared before $t$ and differing by a word from the considered arrival quote; this detection process would include all possible substitutions detected by other models, but would also include many false-positives).

>>>>

In order to start bridging this gap, we set out to \emph{empirically} study public representation transformations at the microscopic level, aiming to stay compatible with macroscopic-level studies of these public representations.
Quotations appeared to be a perfect candidate as public representations.
First, they are usually cleanly delimited by quotation marks (and often with HTML markup in web pages), which greatly facilitates their detection in text corpora.
Second, they stem from a unique ``original'' version, and could ideally be traceable back to that version.
Third, and most importantly, their duplication should \emph{a priori} be highly faithful, apart from cases of cropping: not only should transformations be of moderate magnitude, but when specific words are not perfectly duplicated, it is safe to assume that the variation is due to involuntary cognitive bias --- as writers may expect any casual reader to easily verify, and thus criticize, the fidelity to the original quotation.
Quotation evolution is therefore a perfect environment to measure cognition-induced transformations and relate those findings to macroscopic social dynamics.

@@

The data we use presents an additional challenge: each set of quotations bears no explicit information either about the authoritative original quotation, or about the source quotation(s) each author inspired himself from when creating a new post and reproducing (possibly altering) those sources.
Substitutions are therefore not explicitly encoded in the dataset.

We face a reverse engineering problem where, given all quotations and their occurrence timestamps, we must estimate which was the originating quotation for each instance of each quotation.
We therefore model the underlying quotation selection process by making a few additional assumptions which let us define substitutions from the available data.

The main question in this problem concerns the way we consider later occurrences of a quotation which, when it first appeared, was identified to be an alteration from an original quotation.
Let us give an example: say the quotation ``These accusations are false and absurd'' ($q_a$) appears in a blog on January 18, and the slightly different quotation ``These accusations are false and incoherent'' ($q_b$) appears in another blog on the 19th, the 20th, and the 21st of January.
If $q_a$ was sufficiently prominent when $q_b$ first appeared, we can safely assume that the author of $q_b$ on the 19th based himself on $q_a$.
But what about the following occurrences of $q_b$?
Should we consider them to be substitutions based on $q_a$ (i.e. re-creations of $q_b$ by a new instance of the substitution process that brought from $q_a$ to $q_b$ in the first place) or reproductions of the first occurrence of $q_b$?

To settle this question we model the process as follows: we assume that when a quotation $q$ appears at time $t$, if it is not original (i.e. if not stemming from a source external to the dataset, e.g. initiating a new set of quotations), then it is based solely on the most frequent quotation $q_{max}$ in the preceding period of time $[t - \Delta t ; t]$.
The length $\Delta t$ of that period of time is fixed as a fraction of the total duration of the set of quotations; we took one fifth in the implementation (i.e. one month for a set of quotations spanning five months, or one day for a set spanning only five days; this takes the dynamism of the set of quotations into account).
If $q$ differs from $q_{max}$ by only a word, it is counted as a substitution from $q_{max}$ to $q$. In any other case, i.e. if $q$ and $q_{max}$ are the same or if $q$ and $q_{max}$ are different in other ways, the occurrence of $q$ is not considered to be an instance of substitution and is discarded.

@@

The assumptions embedded in this model are only a subset of a wider set of possibilities, each leading to alternative models.
We identified and implemented five other such models, and they all yielded essentially the same results.
These models differ in the definition of the source quote from which new occurrences stem, essentially modifying the balance between reproduction of previous occurrences of a quote and re-creation of itself by a new substitution instance.
For example, the time-windows considered can have different lengths, can include all occurrences from the beginning of the set of quotations, or can have fixed positions.
The source quotation of a potential substitution can be chosen among those time-windows, or otherwise (e.g. among \emph{all} quotations having appeared before $t$ and differing by a word from the arrival quotation; this detection process would include all possible substitutions detected by other models, but would also include many false positives).
